
---
title: "Reproducible Research Peer Assessment 1"
date: "Sunday, July 19, 2015"
output: html_document

---




### This is the submission of the Assignment 1 for teh Coursera Course of Reproducible Research. Total of 4 questions are answered preceeded by importing the data.


## Pre - Processing

- First of all, as a pre-process, we'll download the file and extract the csv file and create the initail data frame .


```{r "Download"}
setInternet2(use = TRUE)
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", dest="dataset.zip", mode="wb") 
unzip("dataset.zip")
ip <- read.csv("activity.csv",stringsAsFactors=F)
```

- Next we'll pre-process the data for analysing the answers. 


- Below we are pre-processing to answer Question 1. The output dataset is q1prc which calculates total number of steps taken a day.


``` {r PreProcess Q1}
q1prc <- aggregate(ip$steps, by=list(date=ip$date), FUN=sum,na.rm = TRUE)
head(q1prc)
```


- We'll now pre process data for Question 2. Output dataset is q2prc which takes average steps per interval across the dates.


``` {r PreProcess Q2}
q2prc <- aggregate(ip$steps, by=list(interval=ip$interval), FUN=mean,na.rm = TRUE)
head(q2prc)
```

- Pre-Process steps for Q3 and Q4 are done along with their respective sections.



## Question 1: What is mean total number of steps taken per day?


- Below data shows steps taken per day. All 61 day's data (unless fully filled with missing values) are shown.
```{r "Question 1 P1"}
q1prc
```


- Here is a histogran for number of Steps taken per day. 
```{r "Question 1 P2"}
hist(q1prc$x,col = 'blue',main = '# Steps per day', xlab = '# Steps')
```


- Below data provides the mean and median Number of steps taken per day. 
```{r "Question 1 P3"}
mean(q1prc$x)
median(q1prc$x)
``` 



## Question 2: What is the average daily activity pattern?


- To answer the question, we take the intervals and take the average step taken per day in this interval and plot it.
```{r "Question 2 P1"}
plot(q2prc$interval, q2prc$x,type='n',main = "", xlab="", ylab="Average Daily Activity Pattern")
lines(q2prc$interval, q2prc$x,type='l')
```


- Finally we figure out the interval for which average steps taken per day is maximum. Also providing the maximum of Average number of steps with the interval said.
```{r "Question 2 P2"}
q2prc[which(q2prc$x == max(q2prc$x)),]
```



## Question 3: Imputing missing values


- Following calculates number of missings in the dataset.
```{r "Question 3 P1"}
length(which(is.na(ip$steps)))
```


- We take rounded off (to integer) average steps in an interval to fill up the NAs. q3final is the dataset equivalent to the original, with missing values imputed.


```{r "Question 3 P2"}
q3proc <- merge(ip,q2prc,by = "interval")
q3proc <- q3proc[order(q3proc$date,q3proc$interval),]
q3proc$x <- as.integer(q3proc$x)
q3proc[which(is.na(q3proc$steps)),]$steps = q3proc[which(is.na(q3proc$steps)),]$x
q3final <- q3proc[,c("steps","date","interval")]
head(q3final)
```


- Some Pre-Processing for finding new behavior after impuding the missing values.
```{r "Question 3 P3"}
q3prc2 <- aggregate(q3final$steps, by=list(date=q3final$date), FUN=sum,na.rm = TRUE)
head(q3prc2)
```


- Here is the Histogram at Question 1 created with missing values imputed, followed by new mean and medians.
```{r "Question 3 P4"}
hist(q3prc2$x,col = 'blue',main = '# Steps per day', xlab = '# Steps')
mean(q3prc2$x)
median(q3prc2$x)
```


- Mean and median is changed by the below values.
```{r "Question 3 P5"}
abs(mean(q3prc2$x) - mean(q1prc$x))
abs(median(q3prc2$x) - median(q1prc$x))
```



## Question 4: Are there differences in activity patterns between weekdays and weekends?


To answer this quetion we use the data after impuding missing values. 


- First of all, as pre processing, we separate Weekdays and Weekends as different daatasets and take their means separately.
```{r "Question 4 P1"}
q4prc <- q3final
q4prc$day <- weekdays(as.POSIXct(q4prc$date))
q4prcwe <- q4prc[which(q4prc$day == "Sunday" | q4prc$day == "Saturday"),]
head(q4prcwe)
q4prcwd <- q4prc[which(!(q4prc$day == "Sunday" | q4prc$day == "Saturday")),]
head(q4prcwd)
q4prcweag <- aggregate(q4prcwe$steps, by=list(interval=q4prcwe$interval), FUN=mean,na.rm = TRUE)
head(q4prcweag)
q4prcwdag <- aggregate(q4prcwd$steps, by=list(interval=q4prcwd$interval), FUN=mean,na.rm = TRUE)
head(q4prcwdag)
```


- Now we plot the weekends average and weekdays average in 2 different plots.
```{r "Question 4 P2"}
par(mfrow = c(2,1))
plot(q4prcweag$interval,q4prcweag$x,type='l',main = "WeekEnds Average", xlab="", ylab="Steps per Interval") 
plot(q4prcwdag$interval,q4prcwdag$x,type='l',main = "WeekDays Average", xlab="", ylab="Steps per Interval") 

```


- The differences are clerly visible. On weekdays max steps are taken during 8:25-8:30 Morning time, and the rest of the day has very low number of steps on average. whearas for Weekends, the number of steps look to ve distributed over a longer peiod.




## This is the End of this assignment.



#                           :)  Thanks! :)
